---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo=TRUE, message=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
unzip("activity.zip")
activityData = read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
dailySteps = group_by(activityData, date = as.Date(date)) %>% 
    summarise(totalSteps = sum(steps, na.rm = TRUE), .groups="drop")
plot = ggplot(dailySteps, aes(totalSteps)) + geom_histogram(binwidth = 1000)
plot(plot)
```

## What is the average daily activity pattern?
```{r, echo=TRUE}
mean = mean(dailySteps$totalSteps)
median = median(dailySteps$totalSteps)

plot = plot + geom_vline(xintercept = mean, col="red", lwd=1) +
    annotate(geom="text", x=15000, y=8, label="Mean", col="red") +
    geom_vline(xintercept = median, col="blue", lwd=1) +
    annotate(geom="text", x=15000, y=7.5, label="Median", col="blue")
plot(plot)
```

## What is the average daily activity pattern?

```{r echo=TRUE, message=FALSE}
avgInterval = activityData %>% na.omit() %>% group_by(interval) %>%
    summarise(tsteps = mean(steps))

maxInterval = avgInterval[which.max(avgInterval$tsteps),]$interval

avgInterval[which.max(avgInterval$tsteps),]

ggplot(avgInterval, aes(interval, tsteps)) + geom_line() +
  geom_vline(xintercept = maxInterval, col="red", lwd=1)

```

## Imputing missing values

```{r, echo=TRUE, message=F}
completeCases = activityData[complete.cases(activityData$steps), ]
naCases = nrow(activityData) - nrow(completeCases)
naCases #Number of NA cases

## Replace all NA cases with the mean of the respective interval
byInterval = group_by(activityData, interval) %>% summarise(avg= mean(steps, na.rm = T))
activityDataNoNA = activityData;
for(i in 1:nrow(activityData)) {
  if (is.na(activityData[i, 1])) {
      interval = activityData[i, 3]
      meanInterval = byInterval[which(byInterval$interval == interval),]$avg
      activityDataNoNA[i, 1] = meanInterval
  }
}
dailySteps = group_by(activityDataNoNA, date = as.Date(date)) %>% 
    summarise(totalSteps = sum(steps, na.rm = TRUE), .groups="drop")
plot = ggplot(dailySteps, aes(totalSteps)) + geom_histogram(binwidth = 1000)
plot(plot)

```
The values do differ, there are simply more observations so the count for each number of steps increased. The shape of the data remains almost identical. 


3
## Are there differences in activity patterns between weekdays and weekends?

```{r, message=F}
activityDataNoNA$day_type = factor(ifelse(wday(activityDataNoNA$date)== 1 | wday(activityDataNoNA$date)==7, 0, 1), labels=c("weekend", "weekday"))

weekendData = activityDataNoNA[activityDataNoNA$day_type=="weekend",]
weekdayData = activityDataNoNA[activityDataNoNA$day_type=="weekday",]

avgIntervalFinal = activityDataNoNA %>% na.omit() %>% group_by(day_type, interval) %>% summarise(tsteps=mean(steps))


ggplot(avgIntervalFinal, aes(interval, tsteps)) + geom_line() + 
  facet_grid(day_type ~ .) + 
  labs(x="Interval", y="Number of Steps")


```
